---
title: "Data Manipulation"
output: github_document
---

This is for data manipulation

```{r}
library(tidyverse)
```

Load the litters and pups datasets.

```{r}
litters_df=
  read_csv ("./data_import_examples/FAS_litters.csv",
  na = c("",".","NA"))

litters_df = janitor::clean_names(litters_df)

litters_df
```
```{r}
pups_df=
  read_csv ("./data_import_examples/FAS_pups.csv",
  skip = 3,
  na = c("",".","NA"))

pups_df = janitor::clean_names(pups_df)

pups_df
```
## Select some variables

```{r}
select(litters_df, group, litter_number) # select from litter data frame, the group and litter number variable 
select(litters_df, group:gd_of_birth)
select(litters_df, -group) #select everything excep group
select(litters_df, starts_with("gd")) #select everything that start with gd
select(litters_df, starts_with("gd"), group)
select(litters_df, GROUP = group, everything()) #rename group variable and keep everythingelse
select(litters_df, LITTER = litter_number, everything())
select(litters_df, -start_with("gd"))
```

a couple of alternatives

```{r}
rename(litters_df, GROUP = group)
relocate(litters_df, litter_number, gd0_weight)
```

In the pups data, select litter number, sex and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## filter some rows

```{r}
filter(litters_df, gd_of_birth == 19)
filter (litters_df, pups_born_alive <= 5)
filter (litters_df, pups_born_alive != 6)
filter (litters_df, group %in% c("Con7", "Con8"))
filter (litters_df, group %in% c("Con7", "Con8"), pups_born_alive <7)
filter (litters_df, pups_born_alive <4 | pups_born_alive > 8)
drop_na(litters_df, gd0_weight)
```


In the pups dataset, keep only rows where PD walk is less than 11 an sex is 2
```{r}
filter(pups_df, pd_walkM11, sex ==2)
```

## mutate to change some columns 

```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2, 
       group = str_to_lower(group)) 
```
In the pups dataset, create a variable that substracts 7 from PD walk


```{r}
mutate(pups_df, 
       pd_walk_minus7 = pd_walk - 7)
```

## arrange the data

```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))
```






